---
title: "livecode"
author: "Kaitlyn"
date: "3/10/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(mapview)
library(dplyr)
library(measurements)
library(readr)
```


```{r}
fire_perims <- read_sf("data/fire_perims/mtbs_perims_DD.shp")
head(fire_perims)
mapview(fire_perims)
```

If you just want to pull the attributes from the data frame, you can drop the geometry. Then you can make a csv, etc.
```{r}
fire_df <- data.frame(fire_perims) %>% 
  select(-geometry)
```

Extract attributes from the data
```{r}
# can see that you have four kinds of fire
distinct(fire_df, Fire_Type)

# if you want to only take a given type of fire, can filter the sf object using dplyr
fire_perims <- filter(fire_perims, Fire_Type != "RX")
```

## Choosing a coordinate reference system
```{r choose crs}
# check the current 
st_crs(fire_perims) 

st_is_longlat(fire_perims) # checks to see if it is longlat - yes it is, you can't do distance calculations with them

# project it so that it 
fire_proj <- st_transform(fire_perims, crs=3310) # transform into California Albers projection

# check the CRS
st_crs(fire_proj)
```

## Measure area of polygons
```{r measure area}
# function st_area will calculate the area of the polygon
fire_proj <- mutate(fire_proj, area = st_area(fire_proj))

head(fire_proj)

# we already had a column for acres; convert the newly-calculated area in meters-squared back into acres
fire_proj <- mutate(fire_proj, area = measurements::conv_unit(area, from = "m2", to = "acre"))
head(fire_proj) # annoyingly, we still have a note that it is in m^2; it's a named number type, and we want it as a numeric

fire_proj <- mutate(fire_proj, area = as.numeric(area))
```

Sometimes we have spatial information in a non-spatial data format. We will import some data on federally managed campgrounds from [Recreation.gov](https://recreation.gov).

```{r campground data import}
campgrounds <- read_csv("data/recreation/campgrounds.csv")
head(campgrounds)
```

```{r convert csv to spatial}
campground_sf <- st_as_sf(campgrounds, 
                          coords = c("facility_longitude", "facility_latitude"), 
                          crs = 4269, # NAD83
                          remove = F) # won't remove the latitude and longitude columns
mapview(campground_sf)

# let's just subset it to national parks in california
ca_nps <- campground_sf %>% 
  filter(state_code == "CA" & agency_name == "NPS")

mapview(ca_nps)

write_sf(ca_nps, "output/ca_nps.shp")
```

## Drawing buffers
```{r draw buffers}
# project
ca_nps_proj <- st_transform(ca_nps, crs = 3310)

ca_nps_buffer <- st_buffer(ca_nps_proj, 30000)

mapview(ca_nps_buffer)
```

## Intersections

How many national park campgrounds were within 30 km of a wildfire in 2010-2016?
Can see if your 30 km buffers overlap with the wildfire data.

```{r}
mapview(ca_nps_buffer) + 
  mapview(fire_proj, col.regions = "orange")
```

Let's see what intersects. This creates a list object that has the index of all of the ca_nps_buffers that match up with the index of fires - not really ideal for working with.
```{r}
fire_intersects <- 
  st_intersects(ca_nps_buffer, fire_proj)

# convert to a matrix so it's easier to work with
fire_intersects_matrix <- as.matrix(fire_intersects)

# or a key-value format
fire_intersects_keyvalue <- data.frame(fire_intersects)
head(fire_intersects_keyvalue)
```


## Measuring distance

We can measure the distance of our spatial objects from each other. For each of our 61 campgrounds, how far were they from each of the 247 fires? We will return a matrix of dimensions 61 by 247 using `sf::st_distance`. Since our projected data are in meters, the answer will be in meters.

```{r distance}
fire_dist <- st_distance(ca_nps_proj, fire_proj) %>% data.frame()
head(fire_dist)

# coerce to numeric format
fire_dist[] <- lapply(fire_dist, as.numeric)
head(fire_dist)
```

## Testing if objects contain one another

```{r read in smoke data}
# read in NOAA smoke data
smoke <- read_sf("data/smoke/hms_smoke20150819.shp", 
                 crs = 4269) # NAD83

# restrict fires to only those that were burning in sumemr 2015, so they match up with smoke data
fire_2015 <- fire_perims %>% 
  filter(Year == 2015 & StartMonth %in% 6:8)

# mapview
mapview(ca_nps) + 
  mapview(smoke, col.regions = "lightgrey") +
  mapview(fire_2015, col.regions = "orange")
```

```{r testing within}
smoke_proj <- st_transform(smoke, crs = 3310)

# test if campgrounds are within smoke
campgrounds_within <- st_within(ca_nps_proj, smoke_proj)
```

## Bonus - creating a static map

```{r}
ca <- read_sf("data/state/california2016.shp")

ggplot() +
  geom_sf(data=ca) +
  geom_sf(data=ca_nps, color = "darkgreen") +
  geom_sf(data=fire_perims, color = "orange", fill = "orange", alpha = 0.5)

# if you want a projected version of california...

ca_proj <- st_transform(ca, crs = 3310)

ggplot() +
  geom_sf(data=ca_proj) +
  geom_sf(data=ca_nps, color = "darkgreen") +
  geom_sf(data=fire_perims, color = "orange", fill = "orange", alpha = 0.5) +
  theme_bw()
```

